### 2025-03-29 01-Paper/Paper.qmd
Abstract and Intro update

---
title: ""
subtitle: ""
author: 
  - Haowei Fan
thanks: "Code and data are available at: "
date: today
date-format: long
abstract: ""
format: pdf
toc-depth: 2    
number-sections: true
toc: true           
toc-title: "Table of Contents" 
bibliography: references.bib
---

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#### Workspace setup ####
set.seed(912)
# List of necessary packages
packages <- c("tidyverse", "ggplot2", "knitr", "gridExtra", "tidyverse", "lubridate", "forecast", "tseries", "ggplot2", "xgboost", "caret", "slider", "rugarch")
# Install missing packages
missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
install.packages(missing_packages)
# Load packages
lapply(packages, library, character.only = TRUE)

#### Read data ####
prices <- read.csv(here::here("00-data/price.csv"))
climate <- read.csv(here::here("00-data/climate.csv"))
```

# Abstract

This paper forecasts monthly cocoa prices using data from the International Cocoa Organization and climate records from Ghana. After standardized the data, a series of time series models were applied to forecast the prices, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, and GARCH models. 【结果what was found, and why this matters】

# Introduction
Forecasting commodity prices is a challenge in economics and statistical modelling because of the multi-factorial drivers of price behaviour. However, forecasting commodity prices is important for producers, traders and policymakers to have more understanding about the market. Cocoa is a globally traded commodity with significant economic relevance, particularly in regions where it is both produced and consumed at scale such as Ghana. For stakeholders such as producers, traders, and policymakers, accurate forecasting is vital to design procurement strategies, manage supply chain risks, and stabilize income.

Some real-world examples also shows the importance of forecasting the Cocoa price. In 2016–2017, global cocoa prices declined by over 30% [@icco2016november], leading to significant income losses for smallholder farmers in Ghana. However, cocoa exports constitute a major share of national revenue in Chana. This forced some farmers to abandon cocoa cultivation or turn to alternative livelihoods, including environmentally damaging activities such as illegal mining [@bryant2021political]. In order to stabilize the price of cocoa, Ghana and Côte d’Ivoire jointly introduced the Living Income Differential (LID) in 2019, establishing a $400-per-ton premium on cocoa exports to support farmer incomes [@jrc2021lid]. This real-world example shows how price instability can widely influence the economic and social consequences, and highlights the importance of forecasting models.

This paper aims to develop an accurate model for predicting cocoa prices. The paper investigates the monthly behaviour of cocoa prices by using two key datasets, including daily cocoa futures prices from the International Cocoa Organization and daily 	climate data from Ghana, the largest cocoa-producing country in the world. The analysis focuses on modelling the monthly change in log-transformed cocoa prices. The differencing method was use to address non-stationarity. A series of forecasting models was evaluated, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, and GARCH models. Each model was trained on a 70% subsample and assessed using the remaining 30% sample, which is a 70/30 train-test split. Forecast accuracy was assessed with root mean square error (RMSE), AIC, and BIC, with all predictions back-transformed to the original price scale.

结果和为什么重要

# Methodology

###注意，我在此部分写的代码，是为了方便分析，后面全部要删掉。但是后面部分的兄弟可以进行参考###

首先，我们通过previous study确认了几个对于cocoa价格非常重要的几个因素。并且将数据整理成了以月为单位而不是天，每个月的数据是由那个月的平均计算得出。
```{r}
### Load and Aggregate Cocoa Price Data to Monthly
prices <- read.csv(here::here("00-data/price.csv"))
prices$Date <- as.Date(prices$Date, format='%d/%m/%Y')
prices$Price <- as.numeric(gsub(",", "", prices$ICCO.daily.price..US..tonne.))
prices <- prices %>%
  mutate(YearMonth = floor_date(Date, "month")) %>%
  group_by(YearMonth) %>%
  summarise(Price = mean(Price, na.rm = TRUE)) %>%
  ungroup()

### Load and Aggregate Ghana Weather Data to Monthly
climate <- read.csv(here::here("00-data/climate.csv"))
climate$DATE <- as.Date(climate$DATE)
climate <- climate %>%
  mutate(YearMonth = floor_date(DATE, "month")) %>%
  group_by(YearMonth) %>%
  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))

### Merge and Clean Monthly Data
cocoa_data <- left_join(prices, climate, by = "YearMonth")
```

```{r}
### Plot Monthly Time Series
ggplot(cocoa_data, aes(x = YearMonth)) +
  geom_line(aes(y = Price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
```

随后我们给response variable 也就是price 取了一个log，并且做了difference，直到看起来是stationary的。
```{r}
cocoa_data <- cocoa_data %>%
  mutate(log_price = log(Price)) %>%
  drop_na()
### Plot Monthly Time Series
ggplot(cocoa_data, aes(x = YearMonth)) +
  geom_line(aes(y = log_price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "log_price", x = "Date") +
  theme_minimal()
```
```{r}
### Merge and Clean Monthly Data
cocoa_data <- cocoa_data %>%
  mutate(diff_log_price = c(NA, diff(log_price))) %>%
  drop_na()
### Plot Monthly Time Series
ggplot(cocoa_data, aes(x = YearMonth)) +
  geom_line(aes(y = diff_log_price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "diff_log_price", x = "Date") +
  theme_minimal()
```
然后我们数据分成了7:3，其中7成是训练，3成是测试。

我们用训练数据训练了以下的备选模型：ETS，ARIMA，SARIMA，Linear Regression Model，ARCH/GARCH。

模型训练好了以后，我们用测试集测试这些模型的好坏，并且将做了transform的模型，transform predictions back to origin （比如说，做了log，我们就给预测值做一个exponential）

我们通过对比图像以及sum square of erro, AIC，BIC。。。。来对比这些模型。最后选出最好的模型。

# Data
```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
prices <- read.csv(here::here("00-data/price.csv"))
climate <- read.csv(here::here("00-data/climate.csv"))
# Convert date formats
prices$Date <- as.Date(prices$Date, format='%d/%m/%Y')
climate$DATE <- as.Date(climate$DATE)

# Clean price data (remove commas and convert to numeric)
prices$Price <- as.numeric(gsub(",", "", prices$ICCO.daily.price..US..tonne.))

# Select and arrange data
prices <- prices %>% select(Date, Price) %>% arrange(Date)

# Aggregate weather data by date (mean values for daily observations)
climate <- climate %>%
  group_by(DATE) %>%
  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))

# Merge cocoa prices with weather data
cocoa_data <- left_join(prices, climate, by = c("Date" = "DATE"))
cocoa_data <- na.omit(cocoa_data) # Remove missing values

```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-varibales
#| tbl-cap: "Summary Statistics of Important Variables"

tbl_data <- cocoa_data
colnames(tbl_data) <- c("Date", "Price", "Daily Perception", 
                          "Average Temperature", 
                          "Maximum Temperature", 
                          "Minimum Temperature")

knitr::kable(summary(tbl_data))
```

```{r, fig.pos="H"}
#| warning: false
#| message: false
#| echo: false
#| label: fig-series
#| fig-cap: "Time Series of Cocoa Price, Local Average Temperature, and Precipitation"

plot_price <- ggplot(cocoa_data, aes(x = Date, y = Price)) +
  geom_line(color = "blue") +
  labs(title = "Cocoa Price Over Time",
       x = "Date", 
       y = "Cocoa Price in USD") +
  theme_minimal() +
  theme(plot.title = element_text(size = 8),
  axis.title = element_text(size = 8))

plot_temp <- ggplot(cocoa_data, aes(x = Date, y = TAVG)) +
  geom_line(color = "red") +
  labs(title = "Daily Average Temperature Over Time",
       x = "Date", 
       y = "Temperature in °C") +
  theme_minimal() +
  theme(plot.title = element_text(size = 8),
  axis.title = element_text(size = 8))

plot_prcp <- ggplot(cocoa_data, aes(x = Date, y = PRCP)) +
  geom_line(color = "darkgreen") +
  labs(title ="Daily Precipitation Over Time",
       x = "Date", 
       y = "Precipitation in mm") +
  theme_minimal() +
  theme(plot.title = element_text(size = 8),
  axis.title = element_text(size = 8))

grid.arrange(
  plot_price,
  arrangeGrob(plot_temp, plot_prcp, ncol = 2),
  nrow = 2
)
```

```{r, fig.pos="H", fig.width=8,fig.height=5}
#| warning: false
#| message: false
#| echo: false
#| label: fig-pricedecom
#| fig-cap: "Decompose of Series of Prices"

price_ts <- ts(cocoa_data$Price, start = c(year(min(cocoa_data$Date)), month(min(cocoa_data$Date))), frequency = 365)
decomp <- stl(price_ts, s.window = "periodic")

plot(decomp)
```