### 2025-04-01 STA457 Final Project.Rmd
Add files via upload

---
title: "STA457 Final Project"
author: "Wenyi Li"
date: "2025-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
# load required libraries
library(tidyverse)
library(lubridate)
library(forecast)
library(tseries)
library(ggplot2)
library(xgboost)
library(caret)
library(slider)
library(rugarch)
```

# Data Preprocessing
```{r}
# load and preprocess price data
cocoa_prices <- read.csv("Daily Prices_ICCO.csv", stringsAsFactors = FALSE)
cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
cocoa_prices <- cocoa_prices %>%
  mutate(YearMonth = floor_date(Date, "month")) %>%
  group_by(YearMonth) %>%
  summarise(Price = mean(Price, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# load and preprocess Ghana weather data
ghana_weather <- read.csv("Ghana_data.csv", stringsAsFactors = FALSE)
ghana_weather$DATE <- as.Date(ghana_weather$DATE)
ghana_weather <- ghana_weather %>%
  mutate(YearMonth = floor_date(DATE, "month")) %>%
  group_by(YearMonth) %>%
  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
```

```{r}
# Merge and Clean Monthly Data(log + diff)
cocoa_data <- left_join(cocoa_prices, ghana_weather, by = "YearMonth") %>%
  mutate(log_price = log(Price),
         diff_log_price = c(NA, diff(log_price))) %>%
  drop_na()

# Plot Monthly Time Series
ggplot(cocoa_data, aes(x = YearMonth)) +
  geom_line(aes(y = Price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
```

```{r}
# Split Data into Training and Testing Sets(7:3 ratio)
train_size <- floor(0.7 * nrow(cocoa_data))
train_data <- cocoa_data[1:train_size, ]
test_data <- cocoa_data[(train_size + 1):nrow(cocoa_data), ]
```

# Modeling
```{r}
# since not station, already transformed in the pre-processed part(log + diff)
# built ets models
ets_model_1 <- ets(train_data$diff_log_price, model = "ZZZ")
ets_model_2 <- ets(train_data$diff_log_price)
# ets_model_2(ets_model_1) was selected as a candidate
```

```{r}
# built arima, sarima models
# verify stationarity(1st diff + log)
ggplot(train_data, aes(x = YearMonth)) +
  geom_line(aes(y = diff_log_price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
acf(train_data$diff_log_price, main = "ACF of differencing log price")
pacf(train_data$diff_log_price, main = "PACF of differencing log price")
external_regressors <- data.matrix(train_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
arima_model1 <- arima(train_data$diff_log_price, order = c(2,0,2), xreg = external_regressors)
arima_model2 <- arima(train_data$diff_log_price, order = c(2,0,5), xreg = external_regressors)
sarimax_model <- auto.arima(train_data$diff_log_price, xreg = as.matrix(external_regressors), seasonal = TRUE)
summary(arima_model1)
summary(arima_model2)
# arima_model1, sarimax_model was selected as a candidate, seasonality not detected(sarimaxmodel same as arimamodel1)
```

```{r}
# residual diagnostic for arima, sarima models
tsdiag(arima_model1, gof.lag = 20)
tsdiag(arima_model2, gof.lag = 20)
tsdiag(sarimax_model, gof.lag = 20)
```

```{r}
# forecasting in diff-log base
test_xreg <- data.matrix(test_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
test_xreg <- as.matrix(test_xreg)
ets_forecast_2 <- forecast(ets_model_2, h = nrow(test_data))
sarimax_forecast <- forecast(sarimax_model, xreg = test_xreg, h = nrow(test_data))
pred1 <- predict(arima_model1, n.ahead = h, newxreg = test_xreg)
# acc comparison
ets_accuracy_2 <- accuracy(ets_forecast_2, test_data$diff_log_price)
arima1_forecast_vals <- pred1$pred
arima1_accuracy <- accuracy(arima1_forecast_vals, test_data$diff_log_price)
sarimax_accuracy <- accuracy(sarimax_forecast, test_data$diff_log_price)
print("ETS Model Performance:")
print(ets_accuracy_2)
print("ARIMAX Model Performance:")
print(arima1_accuracy)
print("SARIMAX Model Performance:")
print(sarimax_accuracy)
```
```{r}
# back-transform forecasted values
reconstruct_log_prices <- function(last_log_price, diffs) {cumsum(c(last_log_price, diffs))[-1]}

last_log_price <- tail(train_data$log_price, 1)
n <- nrow(test_data)
forecast_dates <- test_data$YearMonth

ets2_log_forecast <- reconstruct_log_prices(last_log_price, ets_forecast_2$mean)
sarimax_log_forecast <- reconstruct_log_prices(last_log_price, sarimax_forecast$mean)
arima_log_forecast <- reconstruct_log_prices(last_log_price, pred1$pred)

ets2_price_forecast <- exp(ets2_log_forecast)
ets2_price_forecast
sarimax_price_forecast <- exp(sarimax_log_forecast)
sarimax_price_forecast
arima_price_forecast <- exp(arima_log_forecast)
arima_price_forecast

forecast_df <- bind_rows(
  tibble(Date = forecast_dates, Forecast = ets2_price_forecast, Model = "ETS Model 2"),
  tibble(Date = forecast_dates, Forecast = sarimax_price_forecast, Model = "SARIMAX"),
  tibble(Date = forecast_dates, Forecast = arima_price_forecast, Model = "ARIMA")
) %>% drop_na()

forecast_df
#forecast_df is a knitted back-transformed forecasted values for three models

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.3) +
  geom_line(data = forecast_df, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.3) +
  labs(title = "Monthly Forecasts vs Actual Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c(
    "ETS Model 2" = "green",
    "SARIMAX" = "blue",
    "ARIMA" = "red"
  )) +
  scale_linetype_manual(values = c(
    "ETS Model 2" = "solid",
    "SARIMAX" = "dotdash",
    "ARIMA" = "twodash"
  ))
```

```{r}
# built linear regression model
# create Lag Features
generate_lags <- function(data, lags = 1:7) {
  for (lag in lags) {
    data[[paste0("lag_", lag)]] <- dplyr::lag(data$log_price, lag)
  }
  return(data)
}
cocoa_data_lagged <- generate_lags(cocoa_data) %>% drop_na()
lm_data <- cocoa_data_lagged %>% select(YearMonth, log_price, starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
train_size <- floor(0.7 * nrow(lm_data))
train_lm <- lm_data[1:train_size, ]
test_lm <- lm_data[(train_size + 1):nrow(lm_data), ]

lm_model <- lm(log_price ~ ., data = train_lm %>% select(-YearMonth))
lm_pred_log <- predict(lm_model, newdata = test_lm) # forecasting values on a log base
lm_pred_price <- exp(lm_pred_log) # back-transform forecasted values
lm_results <- tibble(Date = test_lm$YearMonth, Actual = exp(test_lm$log_price),Predicted = lm_pred_price)
# plot model
ggplot(lm_results, aes(x = Date)) +
  geom_line(aes(y = Actual), color = "red") +
  geom_line(aes(y = Predicted), color = "blue") +
  labs(title = "Linear Regression Forecast vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()
```
```{r}
acf(residuals(lm_model))
library(lmtest)
dwtest(lm_model) #p-value > 0.05 good -> no autocorrelation
```

```{r}
# built GARCH model
log_returns <- diff(log(cocoa_data$Price))
log_returns <- na.omit(log_returns)
train_size <- floor(0.7 * length(log_returns))
train_returns <- log_returns[1:train_size]
test_returns <- log_returns[(train_size + 1):length(log_returns)]
test_dates <- cocoa_data$YearMonth[(train_size + 2):(length(log_returns) + 1)]
garch_spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(1, 0), include.mean = TRUE),
  distribution.model = "norm"
)
garch_fit <- ugarchfit(spec = garch_spec, data = train_returns)
garch_forecast <- ugarchforecast(garch_fit, n.ahead = length(test_returns))
predicted_returns <- as.numeric(fitted(garch_forecast))
last_train_price <- cocoa_data$Price[train_size + 1]
forecast_prices <- last_train_price * exp(cumsum(predicted_returns)) # # back-transform forecasted values
garch_df <- tibble(Date = test_dates,Price = forecast_prices)

# plot GARCH Forecast
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black") +
  geom_line(data = garch_df, aes(x = Date, y = Price), color = "red") +
  labs(title = "GARCH Forecast vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()
```

```{r}
residuals_std <- residuals(garch_fit, standardize = TRUE)
acf(residuals_std, main = "ACF of Standardized Residuals")
Box.test(residuals_std, lag = 20, type = "Ljung-Box") #p-value>0.05 is good -> residuals are noncorrelated
qqnorm(residuals_std)
qqline(residuals_std, col = "red")
```