### 2025-04-04 Final code.Rmd
Add files via upload

---
title: "STA457 Final Project"
author: "Wenyi Li"
date: "2025-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
# load required libraries
library(tidyverse)
library(lubridate)
library(forecast)
library(tseries)
library(ggplot2)
library(xgboost)
library(caret)
library(slider)
library(rugarch)
library(car)
```

# Data Preprocessing
```{r}
# load and preprocess price data
cocoa_prices <- read.csv("Daily Prices_ICCO.csv", stringsAsFactors = FALSE)
cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
cocoa_prices <- cocoa_prices %>%
  mutate(YearMonth = floor_date(Date, "month")) %>%
  group_by(YearMonth) %>%
  summarise(Price = mean(Price, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# load and preprocess Ghana weather data
ghana_weather <- read.csv("Ghana_data.csv", stringsAsFactors = FALSE)
ghana_weather$DATE <- as.Date(ghana_weather$DATE)
ghana_weather <- ghana_weather %>%
  mutate(YearMonth = floor_date(DATE, "month")) %>%
  group_by(YearMonth) %>%
  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
```

```{r}
# Merge and Clean Monthly Data(log + diff)
cocoa_data <- left_join(cocoa_prices, ghana_weather, by = "YearMonth") %>%
  mutate(log_price = log(Price),
         diff_log_price = c(NA, diff(log_price))) %>%
  drop_na()

# Plot Monthly Time Series
ggplot(cocoa_data, aes(x = YearMonth)) +
  geom_line(aes(y = Price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
```

```{r}
# Split Data into Training and Testing Sets(7:3 ratio)
train_size <- floor(0.7 * nrow(cocoa_data))
train_data <- cocoa_data[1:train_size, ]
test_data <- cocoa_data[(train_size + 1):nrow(cocoa_data), ]
```

# Modeling
```{r}
# since not station, already transformed in the pre-processed part(log + diff)
# built ets models
ets_model_1 <- ets(train_data$diff_log_price, model = "ZZZ")
ets_model_2 <- ets(train_data$diff_log_price)
plot(ts(fitted(ets_model_2)), col = "red", 
     main = "Actual vs Predicted for diff-log price ets") + 
  lines(ts(train_data$diff_log_price), col = "green")
# ets_model_2(ets_model_1) was selected as a candidate
```

```{r}
# built arima, sarima models
# verify stationarity(1st diff + log)
ggplot(train_data, aes(x = YearMonth)) +
  geom_line(aes(y = diff_log_price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
acf(train_data$diff_log_price, main = "ACF of differencing log price")
pacf(train_data$diff_log_price, main = "PACF of differencing log price")
external_regressors <- data.matrix(train_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
arima_model1 <- arima(train_data$diff_log_price, order = c(2,0,2), xreg = external_regressors)
arima_model2 <- arima(train_data$diff_log_price, order = c(2,0,5), xreg = external_regressors)
sarimax_model <- auto.arima(train_data$diff_log_price, xreg = as.matrix(external_regressors), seasonal = TRUE)
summary(arima_model1)
summary(arima_model2)
# arima_model1, sarimax_model was selected as a candidate, seasonality not detected(sarimaxmodel same as arimamodel1)
plot(ts(fitted(arima_model1)), col = "red", main = "Actual vs Predicted for diff-log price arima") + lines(ts(train_data$diff_log_price), col = "green")
plot(ts(fitted(sarimax_model)), col = "red", main = "Actual vs Predicted for diff-log price sarimax") + lines(ts(train_data$diff_log_price), col = "green")
```

```{r}
# residual diagnostic for arima, sarima models
tsdiag(arima_model1, gof.lag = 20)
tsdiag(arima_model2, gof.lag = 20)
tsdiag(sarimax_model, gof.lag = 20)
```

```{r}
# forecasting in diff-log base
test_xreg <- data.matrix(test_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
test_xreg <- as.matrix(test_xreg)
ets_forecast_2 <- forecast(ets_model_2, h = nrow(test_data))
sarimax_forecast <- forecast(sarimax_model, xreg = test_xreg, h = nrow(test_data))
h <- nrow(test_data)
pred1 <- predict(arima_model1, n.ahead = h, newxreg = test_xreg)
```

```{r}
# back-transform forecasted values
reconstruct_log_prices <- function(last_log_price, diffs) {cumsum(c(last_log_price, diffs))[-1]}

last_log_price <- tail(train_data$log_price, 1)
n <- nrow(test_data)
forecast_dates <- test_data$YearMonth

ets2_log_forecast <- reconstruct_log_prices(last_log_price, ets_forecast_2$mean)
sarimax_log_forecast <- reconstruct_log_prices(last_log_price, sarimax_forecast$mean)
arima_log_forecast <- reconstruct_log_prices(last_log_price, pred1$pred)

ets2_price_forecast <- exp(ets2_log_forecast)
ets2_price_forecast
sarimax_price_forecast <- exp(sarimax_log_forecast)
sarimax_price_forecast
arima_price_forecast <- exp(arima_log_forecast)
arima_price_forecast

forecast_df <- bind_rows(
  tibble(Date = forecast_dates, Forecast = ets2_price_forecast, Model = "ETS Model 2"),
  tibble(Date = forecast_dates, Forecast = sarimax_price_forecast, Model = "SARIMAX"),
  tibble(Date = forecast_dates, Forecast = arima_price_forecast, Model = "ARIMA")
) %>% drop_na()

forecast_df
#forecast_df is a knitted back-transformed forecasted values for three models

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.3) +
  geom_line(data = forecast_df, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.3) +
  labs(title = "Monthly Predicted of Three Models vs Actual Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c(
    "ETS Model 2" = "green",
    "SARIMAX" = "blue",
    "ARIMA" = "red"
  )) + 
  scale_linetype_manual(values = c(
    "ETS Model 2" = "solid",
    "SARIMAX" = "dotdash",
    "ARIMA" = "twodash"
  ))

# acc comparison
actual_prices <- exp(test_data$log_price)
ets2_acc <- accuracy(ets2_price_forecast, actual_prices)
sarimax_acc <- accuracy(sarimax_price_forecast, actual_prices)
arima_acc <- accuracy(arima_price_forecast, actual_prices)

print("ETS Model Performance:")
print(ets2_acc)
print("ARIMAX Model Performance:")
print(arima_acc)
print("SARIMAX Model Performance:")
print(sarimax_acc)
```

```{r}
# built linear regression model
# create Lag Features
generate_lags <- function(data, lags = 1:6) {
  for (lag in lags) {
    data[[paste0("lag_", lag)]] <- dplyr::lag(data$log_price, lag)
  }
  return(data)
}
cocoa_data_lagged <- generate_lags(cocoa_data) %>% drop_na()
lm_data <- cocoa_data_lagged %>%
  select(YearMonth, log_price, starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
train_size <- floor(0.7 * nrow(lm_data))
train_lm <- lm_data[1:train_size, ]
test_lm <- lm_data[(train_size + 1):nrow(lm_data), ]
lm_model <- lm(log_price ~ ., data = train_lm %>% select(-YearMonth))
plot(ts(fitted(lm_model)), col = "red", 
     main = "Actual vs Predicted for log price regression") + 
  lines(ts(train_data$log_price), col = "green")

lm_pred_log <- predict(lm_model, newdata = test_lm)
lm_pred_price <- exp(lm_pred_log)
lm_results <- tibble(
  Date = test_lm$YearMonth,
  Actual = exp(test_lm$log_price),
  Predicted = lm_pred_price
)
### Plot Regression Results
ggplot(lm_results, aes(x = Date)) +
  geom_line(aes(y = Actual), color = "red") +
  geom_line(aes(y = Predicted), color = "blue") +
  labs(title = "Linear Regression Predicted vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()

lm_accuracy <- accuracy(lm_pred_price, exp(test_lm$log_price))
print("Linear Regression Model Performance:")
print(lm_accuracy)
```

```{r}
acf(residuals(lm_model), main = "ACF of Residuals regression")
library(lmtest)
dwtest(lm_model) #p-value > 0.05 good -> no autocorrelation
qqnorm(residuals(lm_model), main = "QQ-Plot of regression residuals")
qqline(residuals(lm_model), col = "red", lwd = 2)
vif(lm_model)
Box.test(residuals(lm_model), lag = 20, type = "Ljung-Box")
```

```{r}
# Calculate log returns
log_re <- diff(log(cocoa_data$Price))
log_re <- na.omit(log_re)
train_size <- floor(0.7 * length(log_re))
train_re <- log_re[1:train_size]
test_re <- log_re[(train_size + 1):length(log_re)]
test_dates <- cocoa_data$YearMonth[(train_size + 2):(length(log_re) + 1)]
# define garch model - using our best arma model p,q with 2,2 and widely used garchorder 1,1
garch_spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(2, 2), include.mean = TRUE),
  distribution.model = "norm"
)
garch_fit <- ugarchfit(spec = garch_spec, data = train_re)
garch_forecast <- ugarchforecast(garch_fit, n.ahead = length(test_re))
predicted_re <- as.numeric(fitted(garch_forecast))
plot(ts(fitted(garch_fit)), col = "red", 
     main = "Actual vs Predicted for diff-log price garch") +
  lines(ts(train_data$diff_log_price), col = "green")

# back-transform
last_train_price <- cocoa_data$Price[train_size + 1]
forecast_prices <- last_train_price * exp(cumsum(predicted_re))

garch_df <- tibble(
  Date = test_dates,
  Price = forecast_prices
)
# plot the model
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black") +
  geom_line(data = garch_df, aes(x = Date, y = Price), color = "pink") +
  labs(title = "GARCH Predicted vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()

garch_accuracy <- accuracy(forecast_prices, exp(test_data$log_price))

print("GARCH Model Performance:")
print(garch_accuracy)
```
```{r}
residuals_std <- residuals(garch_fit, standardize = TRUE)
acf(residuals_std, main = "ACF of Residuals garch")
Box.test(residuals_std, lag = 20, type = "Ljung-Box")

qqnorm(residuals_std, main = "QQ-Plot of ARMA-GARCH")
qqline(residuals_std, col = "red", lwd = 2)
```

```{r}
# built xgboost model
# generate lags
generate_lags <- function(data, lags = 1:6) {
  for (lag in lags) {
    data[[paste0("lag_", lag)]] <- dplyr::lag(data$log_price, lag)
  }
  return(data)
}

cocoa_data_lagged <- cocoa_data %>%
  generate_lags() %>%
  drop_na()

initial_size <- floor(0.7 * nrow(cocoa_data_lagged))
forecast_horizon <- nrow(cocoa_data_lagged) - initial_size

walk_results <- map_dfr(1:forecast_horizon, function(i) {
  train_set <- cocoa_data_lagged[1:(initial_size + i - 1), ]
  test_point <- cocoa_data_lagged[(initial_size + i), ]

  x_train <- train_set %>% select(starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
  y_train <- train_set$log_price
  x_test <- test_point %>% select(starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)

  dtrain <- xgb.DMatrix(as.matrix(x_train), label = y_train)
  dtest <- xgb.DMatrix(as.matrix(x_test))
# fit model
  model <- xgboost(
    data = dtrain,
    nrounds = 600,
    objective = "reg:squarederror",
    verbose = 0,
    eta = 0.05
  )

  pred_log <- predict(model, dtest)
  tibble(
    Date = test_point$YearMonth,
    Actual = exp(test_point$log_price),
    Predicted = exp(pred_log)
  )
})

ggplot(walk_results, aes(x = Date)) +
  geom_line(aes(y = Actual), color = "black", linewidth = 1) +
  geom_line(aes(y = Predicted), color = "purple", linewidth = 1) +
  labs(title = "XGBoost Walk-Forward Forecast vs Actual (Monthly)",
       y = "Price", x = "Date") +
  theme_minimal()

xgb_accuracy <- accuracy(walk_results$Predicted, walk_results$Actual)
print("XGBoost Walk-Forward Accuracy Metrics:")
print(xgb_accuracy)
```

```{r}
# residual analysis of xgboost model
walk_results <- walk_results %>%
  filter(!is.na(Actual), !is.na(Predicted)) %>%
  mutate(Residual = Actual - Predicted)
acf(walk_results$Residual, main = "ACF of Residuals xgboost")
Box.test(walk_results$Residual, lag = 20, type = "Ljung-Box")

x_train_full <- cocoa_data_lagged %>% select(starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
y_train_full <- cocoa_data_lagged$log_price
dtrain_full <- xgb.DMatrix(as.matrix(x_train_full), label = y_train_full)

final_model <- xgboost(
  data = dtrain_full,
  nrounds = 600,
  objective = "reg:squarederror",
  eta = 0.05,
  verbose = 0
)
```

```{r}
h_future <- 120
ets_forecast_future <- forecast(ets_model_2, h = h_future)

last_log_price <- tail(train_data$log_price, 1)
ets_log_future <- reconstruct_log_prices(last_log_price, ets_forecast_future$mean)

# backtransform
ets_price_future <- exp(ets_log_future)

future_dates <- seq.Date(from = max(cocoa_data$YearMonth) + months(1),
                         by = "month", length.out = h_future)

ets_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = ets_price_future,
  Model = "ETS Model 2"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = ets_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by ETS Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
# calculate historical averages from training data
avg_PRCP  <- mean(train_data$PRCP, na.rm = TRUE)
avg_TAVG  <- mean(train_data$TAVG, na.rm = TRUE)
avg_TMAX  <- mean(train_data$TMAX, na.rm = TRUE)
avg_TMIN  <- mean(train_data$TMIN, na.rm = TRUE)

future_xreg <- matrix(rep(c(avg_PRCP, avg_TAVG, avg_TMAX, avg_TMIN),
                          each = h_future), nrow = h_future)

arima_future <- predict(arima_model1, n.ahead = h_future, newxreg = future_xreg)

# back-transform
arima_log_future <- reconstruct_log_prices(last_log_price, arima_future$pred)
arima_price_future <- exp(arima_log_future)

arima_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = arima_price_future,
  Model = "ARIMA"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = arima_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by ARIMA Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
future_lm_preds <- numeric(h_future)

current_lags <- tail(lm_data$log_price, 6)

for(i in 1:h_future) {
  newdata <- as.data.frame(t(c(current_lags, avg_PRCP, avg_TAVG, avg_TMAX, avg_TMIN)))
  colnames(newdata) <- c(paste0("lag_", 1:6), "PRCP", "TAVG", "TMAX", "TMIN")
  pred_log <- predict(lm_model, newdata = newdata)
  future_lm_preds[i] <- pred_log

  current_lags <- c(pred_log, current_lags[1:5])
}

lm_price_future <- exp(future_lm_preds)

lm_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = lm_price_future,
  Model = "Linear Regression"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = lm_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by Linear Regression Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")

write.csv(lm_forecast_df_10yr, "output_file.csv", row.names = FALSE)
```

```{r}
garch_forecast_10yr <- ugarchforecast(garch_fit, n.ahead = 120)
predicted_10yr_returns <- as.numeric(fitted(garch_forecast_10yr))

last_price <- tail(cocoa_data$Price, 1)
future_garch_prices <- last_price * exp(cumsum(predicted_10yr_returns))

garch_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = future_garch_prices,
  Model = "GARCH"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = garch_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by ARMA-GARCH Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
h_future <- 120
future_xgb_preds <- numeric(h_future)

# Get last 6 log prices (for lag_1 to lag_6)
current_lags <- as.numeric(tail(cocoa_data_lagged$log_price, 6))

# Average weather values (for fixed future input)
avg_PRCP <- mean(cocoa_data_lagged$PRCP, na.rm = TRUE)
avg_TAVG <- mean(cocoa_data_lagged$TAVG, na.rm = TRUE)
avg_TMAX <- mean(cocoa_data_lagged$TMAX, na.rm = TRUE)
avg_TMIN <- mean(cocoa_data_lagged$TMIN, na.rm = TRUE)

# Forecast loop
for (i in 1:h_future) {
  # Construct input as data frame
  newdata <- as.data.frame(t(c(current_lags, avg_PRCP, avg_TAVG, avg_TMAX, avg_TMIN)))
  colnames(newdata) <- c(paste0("lag_", 1:6), "PRCP", "TAVG", "TMAX", "TMIN")

  # Convert to matrix, then to DMatrix
  newdata_matrix <- as.matrix(newdata)
  dnew <- xgb.DMatrix(newdata_matrix)

  # Predict log price
  pred_log <- predict(final_model, dnew)

  # Save prediction
  future_xgb_preds[i] <- pred_log

  # Update lags
  current_lags <- c(pred_log, current_lags[1:5])
}

# Convert log-price to price
xgb_price_future <- exp(future_xgb_preds)

# Generate future dates
library(lubridate)
last_date <- max(cocoa_data_lagged$YearMonth)
future_dates <- seq.Date(from = last_date %m+% months(1), by = "month", length.out = h_future)

# Final forecast data frame
xgb_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = xgb_price_future,
  Model = "XGBoost"
)

# Plot
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = xgb_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices (XGBoost)",
       x = "Date", y = "Price") +
  theme_minimal()
```
```{r}
# Combine all forecast data frames
all_forecasts_10yr <- bind_rows(
  ets_forecast_df_10yr,
  arima_forecast_df_10yr,
  lm_forecast_df_10yr,
  garch_forecast_df_10yr,
  xgb_forecast_df_10yr
)

# Plot the forecasts alongside historical prices
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = all_forecasts_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```